---
title: "Overview of the FARS package"
author: "Lim Wee Pynn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the FARS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include = FALSE}
library(fars)
library(dplyr)
library(maps)
```

This package reads and plots Fatality Analysis Reporting System (FARS) data from the National Highway Traffic Safety Administration.

## Package functions

There are three exported functions available to users:

- `fars_read()`
- `fars_summarize_years()`
- `fars_map_state()`

Further we give a short description with examples how to use the functions. For the purposes of these examples we will use data from FARS, years 2013-2015, that can be found in the package directory under `\extdata` folder.

## Read FARS data files

Function reads in data from a csv file. It takes in a character string of a path of a filename to read into R.

```{r fars_read_example}
filename <- system.file("extdata/accident_2013.csv.bz2", package = "fars")
fars_read(filename)
```

## Summarize number of accidents

Function combines all the dataframes obtained from the `fars_read_years` function into one single dataframe, groups the data into each specific 'year' and 'month', summarises by counting the number of data points for each group, transforms and returns the table from the long to wide format according to the 'year' and number of counts for each group variables.

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "fars"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

Function takes in two numbers specifying a specific state and year and plots the location of accidents according to latitude and longtitude on a state map.

```{r fars_map_state_example}
setwd(system.file("extdata", package = "fars"))
fars_map_state(45, 2015)
```

